### 2021-12-01 slides.Rmd
Rmd slides template

---
title: Analysis of CSV File Reading in R
date: STAT 407 Fall 2021
output: 
  beamer_presentation: 
    theme: Boadilla
    color: beaver
    fonttheme: professionalfonts
    keep_tex: yes
    df_print: kable
    highlight: tango
# csl: apa.csl
# bibliography: ref.bib
# aspectratio: 169
fontsize: 11pt
header-includes:
  - \usepackage{xcolor}
  - \definecolor{luc}{RGB}{88, 41, 49}
  - \setbeamercolor{itemize/enumerate body}{fg=luc}
  - \setbeamercolor{title}{fg=luc}
  - \setbeamercolor{frametitle}{fg=luc}
  - \setbeamertemplate{itemize item}{\color{luc}$\blacktriangleright$}
  - \setbeamertemplate{itemize subitem}{\color{luc}$\blacktriangleright$}
  - \setbeamercolor{footnote}{fg=luc}
  - \setbeamercolor{footnote mark}{fg=luc}
  - \setbeamercolor{footlinecolor}{fg=black,bg=lightgray}
  - \setbeamertemplate{navigation symbols}{}
  - \author[Q. Nguyen \& R. Tedesco]{Quang Nguyen* and Robert Tedesco*}
  - \institute[]{* Loyola University Chicago}
  - \setbeamercovered{transparent}
  - \setbeamercolor{block title}{fg=luc}
  - \setbeamercolor{local structure}{fg=luc}
  - \usepackage{bbding}
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      comment = "")
```

#   Outline

*   Background

*   Data

*   Design

*   ANOVA

*   Additional Analysis

*   Discussion

#   Background

*   Statisticians struggle with reading in large dataset (Matthews twitch)

*   `read.csv` is old

*   Old dogs need to learn new tricks

*   There are better and more efficient algorithms (see book)

    *   `rio::import`

    *   `readr::read_csv` (part of `tidyverse`)

    *   `data.table::fread`